### 2018-02-21 thesis/contents/tn_sagit.tex
Updates from ShareLaTeX

\section{Abstract}
\textbf{Introduction:}  Imaging of trigmeinal neuralgia (TN) has demonstrated key diffusivity alterations in the trigeminal nerve but the difficulties of group tractography analysis have previously prevented us from assessing the entire trigeminal pathway. We apply fully-automated group white-matter tractography to now image the trigeminal sensory pathway, and Gaussian Process (GP) classification to pinpoint key white-matter diffusivity changes that maximally differentiate TN in patients and healthy controls.

\textbf{Methods:} We used SAGIT based group merged tractography for 36 sex-matched TN patients (right-sided pain) and 36 controls, examining the following white matter pathways: trigeminal nerve (CN V), pontine decussation, and thalamocortical fibers. GP classifiers were trained by scrolling a moving window over CN V and S1 tractography centroids. FA, GFA, RD, AD, and MD metrics were evaluated for both groups, analysing TN vs. Control and affected vs. unaffected groups. Classifiers that performed at greater-or-equal-to 70\% accuracy were accepted.

\textbf{Results:} Both affected and unaffected trigeminal nerves in TN patients can be differentiated with 80\% accuracy from controls. Affected CN V FA segments that are differentiated from trons have more overlap towards the distal segments in the cistern and gangion, whereas unaffected CN V segment overlaps extend more towards the brainstem. The bilateral middle segment of S1 can be differentiated from controls. Where the affected S1 RD classifier achieved higher accuracy of 87\%.

\textbf{Conclusions:} We  demonstrate the first instance of group-wise merged tractography analysis of TN vs. control and TN white-matter machine learning classification using GP. The method reproduced previous findings with finer-grained regional specificity. The study also revealed specific diffusivity changes in TPT and S1 white matter pathways.

\section{Introduction}
\subsubsection{Trigeminal Neuralgia}
Trigeminal neuralgia (TN) is a debilitating facial neuropathic pain syndrome characterized by paroxysmal and shock-like pain in one or more dermatomes of the three trigeminal nerve (The fifth cranial nerve: CN V) branches. TN manifests most commonly as classic (also called idiopathic TN), or secondary to a range of neuropathology, including tumors such as meningiomas \cite{Cheng2008}, trigeminal schwannomas \cite{Miller2008}, neurocysticercosis \cite{Revuelta1995}, and most notably multiple-sclerosis (MS) \cite{Cruccu2009,VanderMeijs2002,Nick2012}. The pathophysiology of idiopathic TN is thought to be vascular compression at the root of the CN V nerve entry zone (NEZ) \cite{Linn2011,Love2001}, although idiopathic TN can also occur without evidence of vascular compression \cite{Lee2014}. We previously compared TN and MS-TN diffusivity differences in four segments of the CN V at the level of the pons \cite{Chen2016a}, and confirmed that brainstem CN V diffusivity is altered in MS-TN.

@@

\subsubsection{Along-the-tract analysis}
White matter tractography is commonly used for anatomical visualization \cite{Chen2011b}, white matter segmentation \cite{Behrens2003a,Johansen-Berg2005}, and structural connectivity analysis \cite{Cao2013,Wiech2014}. Obtaining diffusivity metrics from tractography involves converting streamline bundles into a volumetric spatial mask, followed by measurement of the mean metric from all or parts of the masked voxels \cite{Concha2005,Fitzsimmons2009}. However this approach discards the orientation information in the streamlines. The simplest form of spatial masking is to identify voxels where at least one streamline has passed through it, which often results in identifying voxels with minimal streamline pass-throughs. This is potentially problematic when considering that an average 3-Tesla HARDI DWI acquisition has a voxel dimension of 2 to 3 mm. In white matter structures such as CN V as well as other small pathways, the structure diameter may be no more than 3mm, and therefore the naive spatial masking approach would result in severe partial volumes that will confound the result.

In this light, it must be considered that a tractography streamline propagation algorithm involves all or parts of the following stages: 1) subsampling of an initial region for seeding \cite{Basser2002,Cote2012}; 2) identifying a propagation direction, based on previously traversed paths \cite{Malcolm2010,Qazi2009,Tournier2010}; 3) smoothing of the traversed paths for output \cite{Tuch2000d}. It is reasonable to consider these algorithms to be an interpolated subsampling scheme of the DWI volume that takes the diffusion-based orientation into account. Therefore, it would be desirable to directly make use of the tractography streamlines for data measurement.

Along-the-tract analysis has been investigated by various authors \cite{Colby2012,ODonnell2009,Wang2015,Yeatman2012}. In these approaches, the bundles were selectively filtered from global tractography \cite{Wang2015,Yeatman2012}, and limited to major white matter bundles such as the corpus callosum, corticospinal tract, longitudinal fasciculus, arcuate fasciculus, and uncinate. The measures were pooled onto tract centroid in each individual subjects, and then aggregated at the group level. The global tractography approach is unintuitive for anatomy-specific tasks such as neurosurgical planning, where the anatomy of interest is clear. A more focused ROI based delineation approach is therefore preferred in these circumstances.

It is important to note that streamline clustering and centroid identification can be considered as independent tasks. Some clustering methods do not provide a best-representative streamline centroid. Centroid identification involves resampling the streamlines into an identical number of points, and then identifying point correspondences via different distance measures. For these approaches, streamlines require similar length, and an assumption that the start and end of the streamlines across subjects is consistent. However, pathological brain tissue may negatively affect tractography delineation. Therefore the geometry assumptions of such approaches may not be satisfied when studying patient populations. Running time and memory constraints also need to be taken into consideration when applying such algorithms at a group level. We have adopted QuickBundles since it is a practical method of tract clustering and centroid identification with linear running time to obtain S1 streamline centroids. \cite{Garyfallidis2012}

The principal advantage of a streamline-based approach is that we can preserve the fiber orientations. White matter metrics in 3D space can then be projected to a lower dimensional centroid for detailed analysis. We have developed the Selective Automated Group Integrated Tractography (SAGIT) framework (Chen et al., 2016) to automate group-wise tractography delineation, which ensures consistent region-of-interest projection, tractography seeding sampling scheme, and tractography parameter tuning. We then merge the tracts into a template space using non-linear deformations, and  perform tractography centroid identification and statistics at the merged group level to ensure anatomical consistency.

@@

@@

\section{Methods}
\subsubsection{Acquisitions}
We studied 36 patients with classic TN (28 females, 8 males; age 60.1$\pm$16 years; right sided pain), and 36 healthy controls (28 females, 8 males, age 42.7$\pm$12 years).  All patients underwent non-invasive Gamma Knife radiosurgery, with imaging prior and six months post treatment. MRI acquisitions including T1 fast-spoiled gradient echo (FSPGR) and diffusion weighted image (DWI) acquisitions. T1 scans were acquired with 1mm slice thickness, in-plane resolution of 0.9375x0.9375mm, slice spacing=1 mm, TE=5.052 ms, TR=11.956 ms, flip angle=20 deg, FOV=240 mm, and matrix=256x256. DWI scans were acquired with 1 B0 scan, 60 gradient directions, 3mm slice thickness and an in-plane resolution of 0.9375x0.9375 mm, b0=1000 s/mm2, TE=88.6 ms, TR=17000 ms, flip angle=90 deg, field of view (FOV)=120 mm, and matrix=128x128.

@@

\subsubsection{Merged Tractography}
All tractography delineations of the same anatomy across all the subjects and groups were merged into a single tractography file for analysis. A software library\footnote{https://github.com/sinkpoint/fascicle} is created by the author to facilitate this process, the steps are as follows: The individual tract files were converted into a SQLite relational database \cite{owens2010sqlite}, and  all points were exported and transformed by applying the subject DWI to T1, and T1 to template transforms using ANTS' applyTransformToPoints command \cite{Avants2009}. The transformed points were then imported back into the  database, and a new template-space tractography model is exported using the transformed point-set, with all its original scalars and subject/group associations preserved. Finally, all template-space tractography models across all the subjects are merged into a single file.

\subsubsection{Along-the-tract Analysis}
The merged tract was post-processed in 3D Slicer to remove erroneous streamlines that confounded measurement. Streamlines removed include remnant CN V streamlines that stray into the cerebellar peduncles or other non-TGN regions, and streamlines that are perpendicular to the nerve in the cisternal space that appear to be partial delineation of vessels or noise. TPT streamlines that did not decussate at the level of the pons, as well as fibers that stray into the contralateral TGN were removed. S1 streamlines that stray medially towards other regions such as the corpus callosum were also removed.

Tractography centroids, which were the average curve best fit the tract bundle, were identified on the merged tracts in template space. CN V centroids were manually defined using 3 control points. A Catmull-Rom spline \cite{DeRose1988} was then defined from these control points using the VTK software library \cite{Schroeder2005}, which resulted in a curve of length 38.91 mm. The spline is then resampled into 30 points that become the centroid definition with segment length of 1.3 mm. Nearest-neighbor point assignment was used to divide the merged tract into 30 cross sections along the centroids. Diffusion metrics from the points were aggregated for each of the cross-sections.  S1 centroids were defined using QuickBundles \cite{Garyfallidis2012} with distance threshold of 30, resulting in centroid length of 62.1 mm and segment length of 2.07 mm.

@@

Diffusivity metrics of affected (right side) and unaffected CN V and S1 pathways were independently compared with those of controls, where control metrics were the average of both left and right sides. The affected and unaffected sides in TN patients were also compared with each other. 
Age was regressed out by using the residuals of the linear regression model from the controls group. The diffusion values were subsequently normalized to zero mean with ranges $[-1, 1]$ for model training. 
Gaussian Process (GP) classification was performed on CN V and S1 nerve segments to identify key segment differences. Multiple moving windows of length between 5 to 30 were used to constraint tract segments, and a Gaussian Process model was trained for each windowed segment. The entire machine learning process was performed using Scikit-learn machine learning software framework.
For each model, the subjects were split into train and validation sets using stratified 10-fold cross-validation. The GP model was initiated with Matern kernel (length scale = 1), and the model fitted using the training set. The validation set was used to measure prediction accuracy, and the mean accuracy of the 10-fold cross-validation was used to assess the windowed segment. Segments accuracy greater than 75\% were accepted, and the segment length with the greatest accuracy was then determined.

\subsection{Results}
\subsubsection{Tractography}
Merged CN V delineations show anatomical consistency, and extended into the ipsilateral TGN. Changes in FA along the streamlines demonstrate clear differences between cistern and brainstem segments, suggesting good agreement in anatomical registration ( Figure \ref{fig:GPfigure1}). 
The TPT projections, when filtered strictly from the TGN to the ipsilateral VC thalamus region, showed distinct bundles of decussating pathways between left and right. The resulting pathway was not symmetrical, where the right TPT pathway diverts and envelopes the left (Figure \ref{fig:GPfigure2}).
S1 streamlines reliably extended into the face region of the S1 cortex (Figure \ref{fig:GPfigure3}).

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{cnv-inferior-view.png}
\caption{Inferior view of the merged CN V delineations.}
\label{fig:GPfigure1}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{view-decussation.png}
\caption{Close up of the TPT decussation. Red: Right TPT; Blue: Left TPT. Note the Right TPT diverts around the left TPT bundle. }
\label{fig:GPfigure2}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{view-left-right.png}
\caption{The complete trigeminal discriminatory sensory pathway. }
\label{fig:GPfigure3}
\end{figure}

\subsubsection{Diffusivity Differences}
\paragraph{CN V}
Both sides were distinguished from controls with classifiers that achieve around 80\% accuracy (Figure \ref{fig:GPfigure4}). 
Classifiers on the right side have performance ranging from 76\% (GFA) to 86\% (FA) accuracy. The left classifiers achieved accuracy ranging from 79\% (MD) to 81\%(FA). The right side shows more overlap of segment ranges between positions 14--24, in the region of the cistern/gangion; while the left side shows segments that covers the entire the nerve, and extends to cover position 0--4 in the brainstem. Overall, the right side features extend further into the distal space, into the cistern and even the ganglion, while the left side extends more into the brainstem. FA consistently achieved the highest accuracy in both sides, with right FA at position 17--23 (cistern/ganglion) and left FA at 2--16 (brainstem/cister).
Comparing left/right, the FA classifier achieved 75\% accuracy, where its segments range from 11 to 19, which is the cisternal segment. The GFA classifier achieved 70\% accuracy in region 9--12. Demonstrating that affected-unaffected diffusivity differences are found in the cistern/REZ (Table \ref{table:CN V}).

@@

\paragraph{Trigeminopontothalamic tract}
Both left and right TPT pathways can be distinguished from controls with greater than 80\% accuracy (Figure \ref{fig:GPfigureTPT}). The right maximum accuracy was 85\% (FA), and minimum accuracy 83\% (RD), while the left maximum accuracy was 89\% (FA) and minimum 71\% (RD). The regions covered both the decussation and the ascending segments of the TPT, with the right decussation showing more proximal coverage. The left-right classifiers achieved maximum accuracy of 98\% (FA) and minimum of 88\% (MD). The left-right FA covered position 4--14, suggesting strong laterality differences in the TPT (Table \ref{table:tpt}).

@@

\paragraph{S1}
Both sides of the S1 face region pathway were able to be distinguished by classifiers with about 85\% accuracy. The left RD classifier achieved max accuracy of 87\% at position 13--20, while the right RD classifier achieved max accuracy of 83\% at psotion 8--26 (Figure \ref{fig:GPfigure5}). Overall the left classifier ranged from 72\% (AD) to 87\% (RD), and the right classifieres ranged from 73\% (GFA) to 83\% (RD) (Table \ref{table:s1}). Both sides differ from controls in the branching region between the thalamus and cortex. Left-right classifiers showed maximum accuracy of 78\% (RD), and minimum of 76\% (FA). The left-right classifiers overlap in the regions 5--10, at the segment where the pathway emerges from the thalamus.

@@

Overall the GP classifiers resulted in higher precision when compared to accuracy. The overlapping regions revealed differences in localizations that may very well have been missed with singular diffusivity metric analysis. This suggests that cross-metric features should be considered in future studies. This would drastically increase the feature complexity, and thereby require subsequent demand in the number of subjects.

>>>>

\section{Abstract}
\textbf{Introduction:}  Imaging of trigeminal neuralgia (TN) has demonstrated key diffusivity alterations in the trigeminal nerve but technical limitations in group tractography analysis have  prevented us from assessing the trigeminal nerve and its connections in the central nervous system. We apply fully-automated group white-matter tractography to image the trigeminal sensory pathway as well as machine learning classification using Gaussian Process (GP) classifier to pinpoint key white-matter diffusivity changes that maximally differentiate TN patients from healthy controls.

\textbf{Methods:} We used SAGIT based group merged tractography for 36 sex-matched TN patients (right-sided pain) and 36 controls, examining the following trigeminal related white matter pathways: trigeminal nerve (CN V), pontine decussation, and thalamocortical fibers. GP classifiers were trained by scrolling a moving window over CN V and S1 tractography centroids. FA, GFA, RD, AD, and MD metrics were evaluated for both groups, analyzing TN vs. control groups and affected vs. unaffected sides. Classifiers that performed at greater-or-equal-to 70\% accuracy were included.

\textbf{Results:} GP classifier consistently demonstrated bilateral trigeminal changes, differentiating them from controls with an accuracy of 80\%, even though the clinical expression of pain was strictly unilateral. Affected and unaffected sides could also be differentiated from each other with an accuracy of 75\%. Assessment of the trigeminal subregions demonstrated that the segments in affected CN V differ from controls from the region of the REZ and distally. Bilateral TPT could be distinguished from controls with at least 85\% accuracy while TPT left-right classification achieved 98\% accuracy. Bilateral S1 could be differentiated from controls, where the affected S1 RD classifier achieved an accuracy of 87\%.

\textbf{Conclusions:} We  demonstrate the first instance of group-wise merged tractography analysis of TN white-matter machine learning classification using GP, and analysis of the complete trigeminal pathways from the pons to S1 cortex. The previous reported findings of bilateral changes in the REZ of TN patients is observed. Our study suggests additional, bilateral changes extend beyond the REZ and pinpoints these changes with greater detail in regional specificity.

\section{Introduction}
\subsubsection{Trigeminal Neuralgia}
Trigeminal neuralgia (TN) is a debilitating facial neuropathic pain syndrome characterized by paroxysmal and shock-like pain in one or more dermatomes of the three trigeminal nerve (The fifth cranial nerve: CN V) branches. TN manifests most commonly as classic (also called idiopathic TN), or secondary to a range of neuropathology, including tumors such as meningiomas \cite{Cheng2008}, trigeminal schwannomas \cite{Miller2008}, neurocysticercosis \cite{Revuelta1995}, and most notably multiple-sclerosis (MS) \cite{Cruccu2009,VanderMeijs2002,Nick2012}. The pathophysiology of classic TN is thought to be vascular compression at the root of the CN V nerve entry zone (NEZ) \cite{Linn2011,Love2001}, although classic TN can also occur without evidence of vascular compression \cite{Lee2014}. We previously compared TN and MS-TN diffusivity differences in four segments of the CN V at the level of the pons \cite{Chen2016a}, and confirmed that brainstem CN V diffusivity is altered in MS-TN.

@@

\subsubsection{Along-the-tract analysis}
White matter tractography is commonly used for anatomical visualization \cite{Chen2011b}, white matter segmentation \cite{Behrens2003a,Johansen-Berg2005}, and structural connectivity analysis \cite{Cao2013,Wiech2014}. Obtaining diffusivity metrics from tractography involves converting streamline bundles into a volumetric spatial mask, followed by measurement of the mean metric from all or parts of the masked voxels \cite{Concha2005,Fitzsimmons2009}. However this approach discards the orientation information in the streamlines. The simplest form of spatial masking is to identify voxels where at least one streamline has passed through it, which often results in identifying voxels with minimal streamline pass-throughs. This is potentially problematic when considering that an average 3-Tesla HARDI DWI acquisition has a voxel dimension of 2 to 3 mm \cite{Neher2015,Wilkins2015}. In white matter structures such as CN V as well as other small pathways, the structure diameter may be no more than 3mm, and therefore the naive spatial masking approach would result in severe partial volumes that will confound the result.

In this light, a tractography streamline propagation algorithm must involve all or parts of the following stages: 1) subsampling of an initial region for seeding \cite{Basser2002,Cote2012}; 2) identifying a propagation direction, based on previously traversed paths \cite{Malcolm2010,Qazi2009,Tournier2010}; 3) smoothing of the traversed paths for output \cite{Tuch2000d}. It is reasonable to consider these algorithms to be an interpolated subsampling scheme of the DWI volume that takes the diffusion-based orientation into account. Therefore, it would be desirable to directly make use of the tractography streamlines for data measurement.

Along-the-tract analysis has been investigated by various authors \cite{Colby2012,ODonnell2009,Wang2015,Yeatman2012}. In these approaches, the bundles were selectively filtered from global tractography \cite{Wang2015,Yeatman2012}, and limited to major white matter bundles such as the corpus callosum, corticospinal tract, longitudinal fasciculus, arcuate fasciculus, and uncinate. The measures were pooled onto tract centroid in each individual subjects, and then aggregated at the group level. The global tractography approach is unintuitive for anatomy-specific tasks such as neurosurgical planning, where the anatomy of interest is predetermined. A more focused ROI based delineation approach is therefore preferred in these circumstances.

It is important to note that streamline clustering and centroid identification can be considered as independent tasks. Some clustering methods do not provide a best-representative streamline centroid. Centroid identification involves resampling the streamlines into an identical number of points, and then identifying point correspondences via different distance measures. For these approaches, streamlines require similar lengths, and an assumption that the start and end of the streamlines across subjects is consistent. However, pathological brain tissue may  affect tractography delineation. Therefore the geometry assumptions of these clustering approaches may not be satisfied when studying patient populations. Computation time and memory constraints also need to be taken into consideration when applying such algorithms at a group level. We have adopted QuickBundles since it is a practical method of tract clustering and centroid identification with linear running time to obtain TPT and S1 streamline centroids. \cite{Garyfallidis2012}

The principal advantage of a streamline-based approach is that we can preserve the geometric orientations of the pathways. White matter metrics in 3D space can then be projected to a lower dimensional centroid for detailed analysis. We have developed the Selective Automated Group Integrated Tractography (SAGIT) framework \cite{Chen2016} to automate group-wise tractography delineation, which ensures consistent region-of-interest projection, tractography seeding sampling scheme, and tractography parameter tuning. We then merge the tracts into a template space using non-linear deformations, and perform tractography centroid identification and statistics at the merged group level to ensure anatomical consistency.

@@

The streamline measurements resembles time-series data, in that it exhibits positional dependency between measurements. Unlike a time-series however, the measurements represent physical locations along a pathway, therefore they exhibit bi-directional dependency. Posing the study as a supervised machine learning problem, the feature vector would be a N dimensional vector representing N measures at each division. The GP classifier models the feature vector as a latent GP model, before squashing it to produce a logistic output. Therefore the GP model is able to take the positional dependency of the measurements into account, and is well suited for streamline-based classification.

@@

\section{Methods}

\subsubsection{Acquisitions}
We studied 36 patients with classic TN (28 females, 8 males; age 60.1$\pm$16 years; right sided pain), and 36 healthy controls (28 females, 8 males, age 42.7$\pm$12 years).  All patients underwent non-invasive Gamma Knife radiosurgery, with imaging prior and six months post treatment. MRI acquisitions including T1 fast-spoiled gradient echo (FSPGR) and diffusion weighted image (DWI) acquisitions. T1 scans were acquired with 1mm slice thickness, in-plane resolution of 0.9375x0.9375mm, slice spacing=1 mm, TE=5.052 ms, TR=11.956 ms, flip angle=20 deg, FOV=240 mm, and matrix=256x256. DWI scans were acquired with 1 B0 scan, 60 gradient directions, 3mm slice thickness and an in-plane resolution of 0.9375x0.9375 mm, b0=1000 s/mm2, TE=88.6 ms, TR=17000 ms, flip angle=90 deg, field of view (FOV)=120 mm, and matrix=128x128.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{thesis/images/tn-gp-sagit/figure-method.pdf}
\caption{Diagram of the processing steps}
\label{fig:GPMethods}
\end{figure}

@@

\subsubsection{Merged Tractography}
All tractography delineations of the same anatomy across all the subjects and groups were merged into a single tractography file for analysis. We have previously created a software library\footnote{https://github.com/sinkpoint/fascicle} to facilitate this process, the steps are as follows: The individual tract files were converted into a SQLite relational database \cite{owens2010sqlite}, and  all points were exported and transformed by applying the subject DWI to T1, and T1 to template transforms using ANTS' applyTransformToPoints command \cite{Avants2009}. The transformed points were then imported back into the  database, and a new template-space tractography model is exported using the transformed point-set, with all its original scalars and subject/group associations preserved. Finally, all template-space tractography models across all the subjects are merged into a single file. 
The merged tract was post-processed in 3D Slicer to remove anatomically erroneous streamlines. These include remnant CN V streamlines that stray into the cerebellar peduncles or other non-TGN regions, and streamlines that are perpendicular to the nerve in the cisternal space that appear to be partial delineation of vessels or noise. TPT streamlines that did not decussate at the level of the pons, as well as fibers that stray into the contralateral TGN were removed. S1 streamlines that stray medially towards other regions such as the corpus callosum were also removed.

\subsubsection{Along-the-tract Analysis}
Tractography centroids, which were the average curve best fit the tract bundle, were identified on the merged tracts in template space. The spline was then resampled into 30 points and became the centroid definition. Nearest-neighbor point assignment was used to divide the merged tract into 30 cross sections along the centroids. Diffusion metrics from the tractography points were averaged for each of the cross-sections. CN V centroids were manually defined using 3 control points. A Catmull-Rom spline \cite{DeRose1988} was then defined from these control points using the VTK software library \cite{Schroeder2005}. TPT and S1 centroids were defined using QuickBundles \cite{Garyfallidis2012} with distance threshold of 30.

@@

Diffusivity metrics of affected (right side) and unaffected CN V, TPT, and S1 pathways were independently compared with those of controls, where control metrics were the average of both left and right sides. The affected and unaffected sides in TN patients were also compared with each other. 
Age was regressed out by using the residuals of the linear regression model from the controls group. The diffusion values were subsequently normalized to zero-mean with ranges $[-1, 1]$ for model training. 
Gaussian Process (GP) classification was performed to identify key segment differences. Multiple moving windows of length between 5 to 30 were used to constraint tract segments, and a Gaussian Process model was trained for each windowed segment. The entire machine learning process was performed using Scikit-learn machine learning software framework.
For each model, the subjects were split into train and validation sets using stratified 10-fold cross-validation. The GP model was initiated with Matern kernel (length scale = 1), and the model fitted using the training set. The validation set was used to measure prediction accuracy, and the mean accuracy of the 10-fold cross-validation was used to assess the windowed segment. Segments accuracy greater than 70\% were accepted, and the segment length with the greatest accuracy was then determined.

\subsection{Results}
\subsubsection{Anatomical Delineations}
Merged CN V delineations show anatomical consistency, and extended into the ipsilateral TGN. Changes in FA along the streamlines demonstrate clear differences between cistern and brainstem segments, suggesting good agreement in anatomical registration (Figure \ref{fig:GPfigure1}). The CN V centroid curve length was determined as 38.91 mm, with the length of each subdivision being 1.3 mm.
The TPT projections from the TGN to the ipsilateral VC thalamus region showed distinct bundles of decussating pathways in both left and right projections. The resulting pathways were not symmetrical, where the right TPT pathway visibly diverts and envelopes the left (Figure \ref{fig:GPfigure2}).
S1 streamlines reliably extended into the face region of the S1 cortex (Figure \ref{fig:GPfigure3}). The S1 centroid curve length was calculated to 62.1 mm and sub-division length was 2.07 mm.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{cnv-inferior-view.png}
\caption{Merged CN V reconstruction of 36 subjects, with the colour scale depicting FA values derived from raw streamlines. The expect increase in FA as the fibers enter the brainstem can be visualized, and denotes good registration between subjects. }
\label{fig:GPfigure1}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{view-decussation.png}
\caption{Focused view of the TPT decussation. Red: Right TPT; Blue: Left TPT. Note the Right TPT diverts around the left TPT bundle. }
\label{fig:GPfigure2}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{view-left-right.png}
\caption{The complete trigeminal discriminatory sensory pathway, FA is depicted by the colour scales (Red: right side; Blue: left side). The S1 pathways projects to the lateral face region of the S1 cortex. }
\label{fig:GPfigure3}
\end{figure}

\subsubsection{Diffusivity Differences}
\paragraph{CN V}
The classifier distinguished bilateral trigeminal nerves (affected and unaffected) from controls as well as affected trigeminal nerve from unaffected nerves.Both sides were distinguished from controls with classifiers that achieved around 80\% accuracy (Figure \ref{fig:GPfigure4}). 
Classifiers on the right side had accuracie ranging from 76\% (GFA) to 86\% (FA). The left classifiers achieved accuracy ranging from 79\% (MD) to 81\% (FA). The right side showed more overlap of segment ranges between positions 14--24, in the region of the cistern/gangion; while the left side showed segments that covered the entire the nerve, and extended to cover position 0--4 in the brainstem. Overall, the right side classifiers extend further into the distal space, into the cistern and even the ganglion, while the left side extended more into the brainstem. FA classifier consistently achieved the highest accuracy in both sides, with right side at position 17--23 (cistern/ganglion) and left FA at 2--16 (brainstem/cister).
The FA classifier achieved 75\% accuracy in differentiating affected versus unaffected nerves, where its segments ranged from 11 to 19, which was the cisternal segment. The GFA classifier achieved 70\% accuracy in region 9--12. Demonstrating that affected-unaffected diffusivity differences were found in the cistern/REZ (Table \ref{table:CN V}).

@@

\paragraph{Trigeminopontothalamic tract}
Both left and right TPT pathways could be distinguished from controls with greater than 80\% accuracy (Figure \ref{fig:GPfigureTPT}). The right maximum accuracy was 85\% (FA), and minimum accuracy 83\% (RD), while the left maximum accuracy was 89\% (FA) and minimum 71\% (RD). The regions covered both the decussation and the ascending segments of the TPT, with the right decussation showing more proximal coverage. The left-right classifiers achieved maximum accuracy of 98\% (FA) and minimum of 88\% (MD). The left-right FA covered position 4--14, suggesting strong laterality differences in the TPT (Table \ref{table:tpt}).

@@

\paragraph{S1}
Both sides of the S1 face region pathway were able to be distinguished by classifiers with about 85\% accuracy. The left RD classifier achieved max accuracy of 87\% at position 13--20, while the right RD classifier achieved max accuracy of 83\% at psotion 8--26 (Figure \ref{fig:GPfigure5}). Overall the left classifier ranged from 72\% (AD) to 87\% (RD), and the right classifiers ranged from 73\% (GFA) to 83\% (RD) (Table \ref{table:s1}). Both sides differ from controls in the branching region between the thalamus and cortex. Left-right classifiers showed maximum accuracy of 78\% (RD), and minimum of 76\% (FA). The left-right classifiers overlap in the regions 5--10, at the segment where the pathway emerges from the thalamus.

@@

Overall the GP classifiers resulted in higher precision when compared to accuracy. The overlapping regions revealed differences in localizations that may very well have been missed with singular diffusivity metric analysis. This suggests that cross-metric features should be considered in future studies. This would drastically increase the feature complexity, and thereby demand substantial increase in the number of subjects.
